---
title: "Model `r model_list$id` Assessment"
author: "Alison Appling"
date: "`r base::format(base::Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(keras)
library(ggplot2)
library(gridExtra)
```

# Overview

Model `r model_list$id`: `r model_list$description`

```{r summarize_network, echo=FALSE}
model_list$network <- unserialize_model(model_list$serialized)
summary(model_list$network)
```

# Fitting trajectory

```{r summarize_history, echo=FALSE}
model_list$history
```

```{r plot_loss, echo=FALSE}
loss_history <- model_list$history %>%
  as_data_frame() %>%
  filter(metric %in% c('loss'), !is.na(value))
g1 <- loss_history %>%
  ggplot2::ggplot(aes(x=epoch, y=value, color=data)) +
  geom_point(size=0.1, alpha=0.5) +
  scale_color_discrete('') +
  scale_y_log10() + ylab('loss') +
  theme_classic()
g2 <- loss_history %>%
  ggplot2::ggplot(aes(x=epoch, y=value, color=data)) +
  geom_point(size=0.1, alpha=0.5) +
  scale_color_discrete('') +
  scale_y_continuous(limits=c(0,2*median(loss_history$value))) + ylab('loss') +
  theme_classic()
gridExtra::grid.arrange(g1, g2, nrow=2)
```

# Prediction error

```{r preds, echo=FALSE}
preds <- bind_rows(
  data_frame(
    split = 'train',
    example = names(dat$train_target),
    glm = dat$train_input[, 'GLMTemp'] * dat$scales[['GLMTemp']] + dat$centers[['GLMTemp']],
    nn = model_list$network %>% predict(dat$train_input) %>% c(),
    obs = dat$train_target),
  data_frame(
    split = 'devel',
    example = names(dat$dev_target),
    glm = dat$dev_input[, 'GLMTemp'] * dat$scales[['GLMTemp']] + dat$centers[['GLMTemp']],
    nn = model_list$network %>% predict(dat$dev_input) %>% c(),
    obs = dat$dev_target)
) %>%
  mutate( # the following makes assumptions that only work for NN (not RNN or SNN) models
    date=as.Date(substring(example, 1, 8), format='%Y%m%d'),
    depth=as.numeric(gsub('[[:digit:]]{8}_', '', gsub('m', '', example)))
  ) %>% 
  gather(model, pred, glm, nn, obs)
```

```{r rmse, echo=FALSE}
rmse <- function(x1, x2) {
  sqrt(mean((x1 - x2)^2))
}
rmses <- tribble(
  ~model, ~split, ~rmse,
  'glm', 'train', rmse(preds %>% filter(model=='glm', split=='train') %>% pull(pred), preds %>% filter(model=='obs', split=='train') %>% pull(pred)),
  'glm', 'devel', rmse(preds %>% filter(model=='glm', split=='devel') %>% pull(pred), preds %>% filter(model=='obs', split=='devel') %>% pull(pred)),
  'nn', 'train', rmse(preds %>% filter(model=='nn', split=='train') %>% pull(pred), preds %>% filter(model=='obs', split=='train') %>% pull(pred)),
  'nn', 'devel', rmse(preds %>% filter(model=='nn', split=='devel') %>% pull(pred), preds %>% filter(model=='obs', split=='devel') %>% pull(pred))
) %>%
  spread(split, rmse)
rmses
```

# Predictions

```{r plot_preds, fig.width=8, fig.height=7, echo=FALSE}
bg <- 'gray95'
oneyear <- as.Date(c('2009-04-01','2010-01-01'))
tile_height <- 1
colors_preds <- scale_fill_gradientn(na.value=bg, colors=c('blue','orange'))
preds %>%
  ggplot(aes(x=date, y=depth, fill=pred)) +
  geom_tile(height=tile_height) +
  colors_preds +
  scale_y_reverse() +
  theme_classic() +
  facet_grid(model ~ .) +
  theme(panel.background = element_rect(fill=bg)) +
  ggtitle('Predictions and observations by date and depth')
```

```{r plot_preds_1year, fig.width=8, fig.height=5, echo=FALSE}
preds %>%
  filter(date >= oneyear[1] & date < oneyear[2]) %>%
  ggplot(aes(x=date, y=depth, fill=pred)) +
  geom_tile(height=tile_height) +
  colors_preds +
  scale_y_reverse() +
  theme_classic() +
  facet_grid(model ~ .) +
  theme(panel.background = element_rect(fill=bg)) +
  ggtitle(sprintf('Predictions and observations by date and depth (%s)', format(oneyear[1], '%Y')))
```

# Residuals

To get the patterns to pop out while not dropping any data, I'm using a color scale with 5 breakpoints that line up to: the minimum residual, the 5% quantile of all residuals, residual=0 degrees, 95% quantile of residuals, and maximum residual.

```{r plot_resids, fig.width=8, fig.height=5, echo=FALSE}
resids <- preds %>%
  spread(model, pred) %>%
  gather(model, pred, glm, nn) %>%
  mutate(resid = pred - obs,
         resid_percentile = ecdf(resid)(resid))
if(quantile(resids$resid, 0.05) > 0 || quantile(resids$resid, 0.95) < 0) {
  stop('absolute residuals color scheme requires that 0 lies between the 5% and 95% quantiles of all residuals')
}
colors_abs_resids <- with(
  list(resid_breaks=c(quantile(resids$resid, probs=c(0, 0.05)), 0, quantile(resids$resid, probs=c(0.95, 1))) %>%
         setNames(., paste(sprintf('%0.1f', .), sprintf('(%s)', c('min', '5%', 'perfect', '95%', 'max'))))), 
  scale_fill_gradientn(
    'Residual',
    na.value=bg, colors=c('#2c7bb6','#abd9e9','#ffffbf','#fdae61','#d7191c'),
    breaks = resid_breaks,
    values = scales::rescale(resid_breaks),
    limits = range(resid_breaks)))
resids %>%
  ggplot(aes(x=date, y=depth, fill=resid)) +
  geom_tile(height=tile_height) +
  scale_y_reverse() +
  colors_abs_resids +
  theme_classic() +
  theme(panel.background = element_rect(fill=bg)) +
  facet_grid(model ~ .) +
  ggtitle('Predictions - Observations (all years)')
```

```{r plot_resids_1year, fig.width=8, fig.height=5, echo=FALSE}
resids %>%
  filter(date >= oneyear[1] & date < oneyear[2]) %>%
  ggplot(aes(x=date, y=depth, fill=resid)) +
  geom_tile(height=tile_height) +
  colors_abs_resids +
  scale_y_reverse() +
  theme_classic() +
  theme(panel.background = element_rect(fill=bg)) + 
  facet_grid(model ~ .) +
  ggtitle(sprintf('Predictions - Observations (%s)', format(oneyear[1], '%Y')))
```

# Predictions vs Observations

```{r plot_preds_pvo, echo=FALSE}
preds %>%
  spread(model, pred) %>%
  gather(model, pred, glm, nn) %>%
  ggplot(aes(x=obs, y=pred, color=model)) +
  geom_point() + 
  geom_abline() +
  scale_color_manual(values=c('#64dd17','#ef6c00')) +
  facet_grid(model ~ split) +
  theme_classic() +
  ggtitle('Predictions vs Observations (all years)')
```
